---
title: "Homework 2"
author: "Alice Tivarovsky"
date: "9/26/2019"
output: github_document
---
# Problem 1

## Mr Trash Wheel Data
We are reading in the Mr. Trash Wheel dataset, stored as an excel file. We will also remove the last column, which contains notes. 

```{r data_frame1}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)

# Importing Mr Trash Data
mr_trash_data = read_excel("../data/mr_trash_dataset.xlsx")
mr_trash_data = janitor::clean_names(mr_trash_data)

# Removing last column of comments
mr_trash_data = select(mr_trash_data, -x15)
```

The "raw" dataframe contains 336 entries and 15 columns. 

We will now remove any rows that have dumpster = NA and convert the column "sports_balls" to an integer vector rounded to the nearest whole number. 

```{r data_frame2}

# Removing rows with "NA" for dumpster
mr_trash_data = drop_na(mr_trash_data, dumpster)

# Rounding sports_balls to nearest integer
mr_trash_data = mutate(mr_trash_data, sports_balls = round(as.integer(sports_balls, digits = 0)))

```

## Precipitatin Data

We will now import and clean 2017 and 2018 precipitation data from the Mr. Trash dataset. 

```{r data_frame3}
# Importing 2017 and 2018 precipitation data
precip_data_18 = read_excel("../data/mr_trash_dataset.xlsx",3)
precip_data_17 = read_excel("../data/mr_trash_dataset.xlsx",4)
precip_data_18 = janitor::clean_names(precip_data_18)
precip_data_17 = janitor::clean_names(precip_data_17)
```

We note that the 2017 dataset is complete, but the 2018 dataset is missing values for August - December. We also note that both datasets contain a row at the end with total precipitation for the year, which we don't need. We will omit these rows, along with the missing values in 2018, and add a year variable before combining the datasets. 

```{r data_frame4}
# Removing rows with "NA" for precipation 
precip_data_17 = drop_na(precip_data_17, precipitation_in)
precip_data_18 = drop_na(precip_data_18, precipitation_in)
precip_data_18 = drop_na(precip_data_18, x2)
```

